---
title: "STAT 442/842, CM 762 W25 Assignment 4"
subtitle: "DUE: Friday April 4 by 11:59pm EST"
output:
  pdf_document:
    keep_tex: yes
    number_sections: no
  html_document:
    toc: yes
  word_document: default
urlcolor: blue
---


## NOTES
NOTES
This assignment has 4 programming questions for everyone.

Your assignment must be submitted by the due date listed at the top of this document, and it must be
submitted electronically in .pdf format via Crowdmark.

This means that your responses for different question parts should begin on separate pages of your .pdf file.
Note that your .pdf solution file must have been generated by R Markdown. 

Additionally: Organization and comprehensibility is part of a full solution. Consequently, points will be deducted for
solutions that are not organized and incomprehensible. Furthermore, if you submit your assignment to
Crowdmark, but you do so incorrectly in any way (e.g., you upload your Question 2 solution in the Question
1 box), you will receive a 5% deduction (i.e., 5% of the assignment’s point total will be deducted from your
point total).

There are 50 marks for each student in the class. There are 0 marks for each student outside of the class.
There are 25 marks for students that are in a quantum superposition of inside and outside the class until
their waveforms collapse, at which point their grades will be updated; if this is a required class, then these
students will be considered to have graduated and not graduated at the same time until they are observed
at a convocation.

\newpage


Question 0 (0 marks), install the tidyr package, run the following code on the chemistry data from Lake
Simcoe and save the results as dat_wide.

A full description of the data comes from here:
https://files.ontario.ca/moe_mapping/downloads/metadata/opendata/Lake_Simcoe_Monitoring_metadata_EN.pdf


The dataset itself comes from the Ontario Data Catalogue here: 
https://data.ontario.ca/en/dataset/lakesimcoe-monitoring

```{r, eval=FALSE}
library(tidyr)
dat = read.csv("SimcoeChemistryData1980-2022.csv")
dat$STATYEAR = paste(dat$STATION, dat$YEAR, sep="-")
dat2 = dat[, c("STATION", "YEAR", "STATYEAR","TEST.VALUE","TEST.CODE")]
dat_wide <- dat2 %>% pivot_wider(names_from = TEST.CODE,
                    values_from = TEST.VALUE, values_fn = mean)


dat_wide = data.frame(dat_wide)


```



\newpage


## Question 1 (15 marks)


Using `datq1` and the following code...

```{r, eval=FALSE}
datq1 = subset(dat_wide, STATION == "K42" & YEAR %in% c(1990,2000,2010,2020))
datq1 = datq1[, c("STATYEAR","PH","NNOTUR","ALKT","SIO3UR","DIC","DOC")]
```

You may use information from the following webpages (and any other sources, but these are recommended)
• https://www.datanovia.com/en/blog/beautiful-radar-chart-in-r-using-fmsb-and-ggplot-packages/
• https://r-charts.com/ranking/ggradar/

a) (10 marks)


Make 4 radar charts with ggradar of PH, NNOTUR, ALKT, SIO3UR, DIC, and DOC for Station K42 for the years
1990, 2000, 2010, and 2020.

• Use facets to label each year
• Give the whole plot the title “Kempenfelt Bay 42”
• Label the variables “pH”, “Nitrogen”, “Alkalinity”, “Silicon”, “Inorganic”, and “Organic” respectively.
• Choose appropriate gridline colours and types.


b) (5 marks)

Make the same graph as in part a, but this time do it all in one graph, with different colours representing
different decades. You may need to manipulate the dataset in order to set up and use a grouping variable.
Make sure you include a legend for the decades.

\newpage

## Question 2 (15 marks)

Use scatter3d in plot_ly in the plotly package and the following code to do the following:

```{r, eval=FALSE}
datq2 = subset(dat_wide, STATION == "K42" & YEAR >= 1986)
datq2 = datq2[, c("YEAR","PH","ALKT","DOC")]
```


Plot a 3D graph such that

• There are data points drawn at PH (x), ALKT (y), and DOC (z) for each year that has complete
information.
• A colour scheme that is selected such that the earliest year is blue, and the latest year is red.
• There are guidelines from each point going down to the z=minimum plane.
• The camera angle is set to show the data points with a minimal amount of overlap, while still showing
all three axes
• The axes are labelled accordingly
• There is an appropriate title.


\newpage


## Question 3 (20 marks)

Using the following code and files.

The shapefile for Lake Simcoe was subsetted from 

https://open.canada.ca/data/en/dataset/448ec403-6635-456b-8ced-d3ac24143add

using the following code. You need both `ghy_000c11a_e.shp`, and `ghy_000c11a_e.shx`, which includes all of the shapefiles of bodies of water in Canada. This code is just included in case you want to do something geographic for the final project.

This code gives you `datq3` which includes the dissolved inorganic solids (DIC), latitude (LAT), and longitude (LONG) of each `STATION` for the `YEAR`s 1990, 2000, 2010, and 2020.



```{r, eval=FALSE}
# Read this shape file with the sf library.
library(sf)
library(stringr)
sf_all <- read_sf("ghy_000c11a_e.shp") #read_sf also looks for the .shx file

dat_sf = subset(sf_all, NAME %in% c("Lake Simcoe", "Cook's Bay", "Kempenfelt Bay"))
sf_all = NULL

dat_stn  <- read.csv("LakeSimcoeStations.csv")
dat_stn$DESCRIPTION = str_replace(dat_stn$DESCRIPTION, " eupho composite", "")

datq3    <- subset(dat_wide, YEAR %in% c(1990,2000,2010,2020))
datq3    <- datq3[,c("STATION","YEAR", "DIC")]
datq3$LAT  <- NA
datq3$LONG <- NA
datq3$DESC <- NA


for(k in 1:nrow(dat_stn))
{
  this_stn = dat_stn$STN[k]
  idx = which(datq3$STATION == this_stn)
  datq3$LAT[idx] = dat_stn$LATITUDE[k]
  datq3$LONG[idx] = dat_stn$LONGITUDE[k]
  datq3$DESC[idx] = dat_stn$DESCRIPTION[k]
}
```


a) (15 marks) Make a plot of Lake Simcoe with the outline of the lake (it's okay to include to boundary lines between Lake Simcoe and Cook's Bay, and between Lake Simcoe and Kempenfelt Bay). Draw data points of each of the stations, and colour code to the levels of DIC. Use appropriate titles, labels, and theming. Make sure the data points are large enough to see colours easily, but not so large that overlapping is a major issue. Facet by year.


```{r}
library(ggplot2)
library(sf)
library(ggspatial)
library(ggmap)
library(ggrepel)
library(ggplot2)


# Plot lake simoe
ggplot() +
  geom_sf(data = dat_sf, aes(fill = NAME), alpha = 0.5) +
  geom_point(data = datq3, aes(x = LONG, y = LAT, colour = DIC), size = 3) +
  scale_fill_manual(values = c("blue", "green", "red")) +
  scale_colour_gradient(low = "blue", high = "red") +
  labs(title = "Lake Simcoe Stations",
       x = "Longitude",
       y = "Latitude",
       fill = "Lake Name",
       colour = "DIC") +
  theme_minimal() +
  facet_wrap(~ YEAR) +
  theme(legend.position = "bottom") +
  coord_fixed()
```


b) (5 marks) Make a plot of Lake Simcoe with the outline of the lake as in part a. This time label the stations with their descriptions. Don't panel by year or colour code the points.
